---
title: "GSS challange Less Data"
author: "Maximiliano Ortiz"
date: "Fall 2025"
output: html_document
---



```{r include=FALSE,echo=FALSE}
source(file="fisherXY.R")
require(tidyverse)
require(tigerstats)
require(knitr)
require(foreign)
GSS <-read.dta(file="GSS2016.dta")
GSS <- droplevels(GSS)
mydata <- select(GSS,sex,race,cappun,degree,relig,gunlaw,tvhours,owngun,age,wordsum)
mydata <- na.omit(mydata)
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

Is attitude toward capitol punishment (cappun) independent of race (race)?
If not, then whose view differs from whom??

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->

```{r}
dd3 <- mydata %>% group_by(race,cappun) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd3,aes(x=race,y=count,fill=cappun))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd3,aes(x=race,y=prcnt*100,fill=cappun)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```
  
The white people are for it with about 2 to 1. The black people are more against it, and others seem more for it.

## Now with LESS data (more variables give less data due to NA's) create the approate graphic displays to answer the question.





### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
table2 <- xtabs(~cappun + race, data=mydata)
rowPerc(table2)
colPerc(table2)
```

The majority of white people favor capital punishment, with 67 .75% supporting it. 79 Black people only had 47.24% and others had a significant value of 61.25%. The majority of black people opposed capital punishment with a value of 52.76%. White people only opposed it by 32.25%, and others only opposed it by 38.75%.

## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
fisherXY(mydata$cappun,mydata$race,printAll = TRUE)
```

You can mostly see that the black and white are the ones with the most difference, according to the p-values. It has a very low lumber of around 0.000017. More white people are for capital punishment, while less black people are for capital punishment. However, between white and s-value is only 0.2565. This means that there really isn't a difference between the amount of people who favor capital punishment for They both prefer capital punishment. Lastly, the p-value between black and others is only 0.06299, This shows that they are also close. It is less than the p-value of white and others, but it is significantly not different. Black people are more against capital punishment than others.


# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->


While this process was similar to the one done in the master branch , there is a significantly different amount of observations. This is due to the fact that we only concluded a statistical difference between the black and white. This only gave us 821 observations. This led to there not being much difference in p-values, except for the black and white. There is less to go off of, when you have less observations. Which was backed up by what I observed during fisher's test in the number2 branch. You can mostly see that the black and white are the ones with the most difference, according to the p-values. It has a very low number of around 0.000017 More white people are for capital punishment, while less black people are for capital punishment. However, between white and other, the p-value is only 0,2565. This means that there really isn't a difference between the amount of people who favor capital punishment for white and others. They both prefer capital punishment. Lastly, the p-value between black and others is only 0.06299. This shows that they are very close. The graphical and numerical descriptive results showed the same thing. They both showed percentage and visual values of the black and white being different because black people are against capital punishment, and white and others being almost the same because they both support capital punishment more.

